---
title: "R Analysis: Penguin Body Mass"
author: "Jackson Odumodu"
date: 2026-09-23
---

## Data

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER. The data were collected by Dr. Kristen Gorman and the Palmer Station LTER program, and packaged for R by Horst, Hill and Gorman (2020). It is released under a CC0 licence.

This post asks a simple question: do the three penguin species differ in body mass?

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(palmerpenguins)
library(knitr)
```

## Cleaning the data

Some penguins are missing a body mass measurement, so drop those rows.

```{r}
penguins_clean <- penguins |>
  drop_na(body_mass_g)

nrow(penguins_clean)
```

## Summary table

```{r}
#| label: tbl-summary
#| tbl-cap: "Body mass by species."

summary_table <- penguins_clean |>
  group_by(species) |>
  summarise(
    n = n(),
    mean_mass_g = round(mean(body_mass_g), 1),
    sd_mass_g = round(sd(body_mass_g), 1)
  )

kable(summary_table)
```

@tbl-summary shows that Gentoo penguins have the highest mean body mass, at `r summary_table$mean_mass_g[summary_table$species == "Gentoo"]` g.

## Figure

```{r}
#| label: fig-penguin-mass
#| fig-cap: "Body mass of penguins by species."

ggplot(penguins_clean, aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(x = "Species", y = "Body mass (g)")
```

## What I found

@fig-penguin-mass agrees with the table. Gentoo penguins are clearly heavier than the other two species, while Adelie and Chinstrap penguins have similar body masses.

